This thesis and my associated research endeavors would not have been possible without the support of so many people, and for that I am deeply thankful.

In particular, I would like to thank my advisor, Prof. John Hansman, for his unwavering support and guidance throughout my time at MIT. I'm also indebted to Prof. Mark Drela, both for his personal mentorship and teaching, as well as his broader professional contributions -- much of the work in this thesis would not have been possible without his prolific contributions to the field of aerospace engineering. I would also like to thank Prof. Karen Willcox, who graciously agreed to serve on my doctoral committee and has provided wise and valuable insights throughout our meetings. I am also grateful for Dr. Tony Tao and Prof. Joaquim Martins, both for serving as readers on this thesis and for their contributions to the field.

I would like to thank the open-source scientific computing community for their contributions to this work. In particular, I'd like to thank the maintainers of NumPy, SciPy, CasADi, JAX, IPOPT, and the Python programming language for their tireless work on these tools that have enabled so much of the research in this thesis. The modern day open-source software ecosystem epitomizes the spirit of worldwide collaboration that Newton described so long ago: ``If I have seen further, it is by standing on the shoulders of giants.'' It is a privilege to contribute to this community.

Various parts of the research in this thesis were funded by the MIT Portugal Program, BAE Systems, the National Defense Science and Engineering Graduate Fellowship, and I thank them for their support. I would also like to thank the MIT Supercloud team for supporting my free access to supercomputing resources that enabled parts of this research.

At MIT AeroAstro, I would like to thank my labmates in ICAT for their camraderie and support over the years, especially Matt Vernacchia, Kelly Mathesius, Julia Gaubatz, Annick Dewald, Ara Mahseridjian, Estelle Martin, Trevor Long, Juju Wang, Sandro Salgueiro, Marek Travnik, Mina Cezairli, and many others. I've been so fortunate to spend the past few years alongside such a talented and friendly group of people.

Also at AeroAstro, I'd like to thank machine shop staff Todd Billings and Dave Robertson for their friendship and help over the years. When you've been stuck on a hard math problem for days, sometimes there's nothing like cutting some balsa and building a model airplane to clear your head.

Closer to home, I'd like to thank my friends and family for their support and encouragement -- it means the world to me. Most of all, I'd like to thank my partner and best friend, Marta Manzin, for her seemingly infinite love and support throughout this journey. She has been my rock through the ups and downs of graduate school. I cannot imagine having done this without you.